Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2018-12-27 Paper/sync.tex
almost 2nd draft

Angular momentum evolution in low-mass (M$\lsim 1$ M$_{\odot}$) short-period (P$_{orb} \lsim 10$ d) stellar binaries is strongly influenced by tides.  Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}. Orbital circularization is ubiquitous for short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized. For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.  Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster M35 tend to have circular orbits.

Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins. Tidal torques drive them towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.  Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008}, and as a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.

The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d. Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking. These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$. Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.

Understanding how tidal torques affect angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g. Praesepe, ][]{Douglas2017} and field stars \citep[e.g. \kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys. In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.  Even if stellar binaries have not yet tidally-locked, tidal torques modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, imparting a contaminating signal that is not currently accounted for by modern models. Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology, a method that links the long-term spin-down of low-mass stars to their ages, \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect.

Here, we examine angular momentum evolution over the full main sequence lifetimes of low-mass stellar binaries using a realistic treatment of stellar evolution, magnetic braking, and tidal torques. We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.  We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.  We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.  We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.

XXX Why not discuss tidal braking more? I expected an intro that described tidal affects on binaries, magnetic braking, and then stating that no study had properly linked the two. The tidal review is good, but more on MB would be helpful. XXX

tides impact binary ecc, tides can brake spins, tidally-lock systems, keeping them fast rotators etc. Talk about magnetic braking as both magnetic braking and tides act in binaries.  Talk about observations of Prot, like  \citet{Abt2004} citation and similar things like Lurie

@@

In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar, using both magnetic braking models. We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr. All stars' radii contract along the pre-main sequence, spinning the stars up (right panel). Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales. Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution. The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.

@@

Here we examine the P$_{rot}$ evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a focus on systems with weak tidal torques, i.e. long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking.  In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms. Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}. The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.

@@

\subsubsection{\citet{Reiners2012}}

The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques. XXX

@@

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr. For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions. Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
    \label{fig:protDist}%
\end{figure*}

@@

XXX

@@

Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g. the right panel of Fig.~\ref{fig:protDist}.  If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal.

In general, itt is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue. Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are subject to large systematic errors.  Moreover, this problem could be more significant that the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread. We caution that any application of gyrochonlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.

@@

Both models can reproduce the major trends in the data, but the predictions of the CPL model fail to reproduce several key aspects.  First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.  This prediction is inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.  The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.  The CTL model, however, seems to excel where the CPL models fails by producing both populations. We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.

@@

XXX The abov discussion is a bit tricky. Remember that CPL and CTL are truncated infinite series, so locking into 1:1 vs 3:2 might not be the only possible outcomes. In Murray and Dermott they show how one can use a CPL model to produce the "pseudo-equlibrium" period that is a continuous function of $e$. XXX

The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model. The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making this distinction conclusive.  More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.

Third, as shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 70$ d would provide strong evidence in favor of the CPL model as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d, e.g. Fig.~\ref{fig:lockedCTL}, whereas the CPL model readily tidally-locks binaries out to P$_{orb} \approx 100$ d.  We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}. Long term spectroscopic monitoring may be warranted in such cases.

>>>>

The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Dunn1961,Mestel1968}. Early on in stellar lifetimes, stars spin-up as they contract along the premain sequence.  Once stars reach the main sequence, stellar radii remain mostly constant and magnetic braking saps angular momentum from low-mass stars, gradually spinning them down over time \citep{Skumanich1972}. Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g. Praesepe, ][]{Matt2015,Douglas2017} and field stars \citep[e.g. \kepler, ][]{Reiners2012,Matt2015,vanSaders2018}. Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronlogy \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.

In contrast, the angular momentum evolution in low-mass short-period (P$_{orb} \lsim 10$ d) stellar binaries is dominated by tides.  Tidal torques drive secular changes in the binary orbit and stellar spins, eventually circularizing the orbit and synchronizing the stellar spins in the long-term \citep{Counselman1973}. Orbital circularization is ubiquitous for short-period binaries, owing to the tidal torque's strong semi-major axis dependence, with both theoretical \citep[e.g.][]{Zahn1989,Claret1995} and observational \citep[e.g.][]{Meibom2005,Mazeh2008,Lurie2017} studies finding that most binaries with P$_{orb} \lsim 10$ d are circularized. For short-period binaries, tidal torques work quickly on ${\sim}100$ Myr timescales, as \citet{Zahn1989} finds that the orbit of solar twin binaries circularizes during the stellar pre-main sequence.  Observations by \citet{Meibom2005} support this picture as they find short-period binaries in the ${\sim}150$ Myr old cluster M35 tend to have circular orbits.

Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins. Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.  Tidal-locking occurs much earlier than orbital circularization as the tidal-locking timescale is estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is less angular momentum in stellar spins than the binary orbit. As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.

In low-mass binaries, both magnetic braking and tidal torques compete to shape the stellar P$_{rot}$ evolution. When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$. In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2017}. Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[e.g. P$_{rot} >$ P$_{orb}$, ][]{Habets1989,Zahn1994,Keppens1997}. This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations.

The influence of tides is not necessarily limited to short P$_{orb}$ systems either, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered. The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.

%The large sample of $816$ P$_{rot}$ measurements derived using starspot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d. Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking. These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$. Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.

Understanding how tidal torques and magnetic braking impact angular momentum evolution in low-mass stellar binaries is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g. Praesepe, ][]{Douglas2017} and field stars \citep[e.g. \kepler, ][]{McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys. In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.  Tidal torques can modify stellar rotation periods towards the tidally-locked state, away from the long-term spin-down due to magnetic braking predicted for single stars \citep{Dunn1961,Skumanich1972,Barnes2003}, or compete against magnetic braking to form subsychronous rotators, imparting a contaminating signal that is not currently accounted for by models. Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect.

There is currently a large number of \kepler binaries with known P$_{rot}$ and P$_{orb}$ \citep[e.g.][]{Lurie2017}. Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} are expected to detect additional low-mass eclipsing binaries, with Gaia parallaxes \citep{Gaia2016} poised to help refine these stellar parameters, potentially creating a rich dataset of the angular momentum budgets of low-mass binaries. In today's data-rich era of astronomy, it is critical to develop a theoretical framework to understand the angular momentum evolution of low-mass binaries due to stellar evolution, magnetic braking, and tidal torques.  Such a framework can enable the characterization of the nature of tidal torques in binaries and allow for the long-term modeling of the evolutionary histories of individual systems.

Here, we present a model for the angular momentum evolution of low-mass stellar binaries over their full premain and main sequence lifetimes using a realistic treatment of stellar evolution, magnetic braking, and tidal torques. We investigate under what conditions tidal-locking occurs, and how tidal torques influence rotation in stellar binaries as a function of binary P$_{orb}$ and tidal dissipation parameters for two widely-used equilibrium tidal models.  We show how tidal torques can impact stellar rotation in binaries out to P$_{orb} = 100$ d, causing stellar rotation periods to not strongly correlate with age, making the predictions of gyrochronlogy models to fail in such systems.  We describe our model in $\S$~\ref{sec:methods} and our simulation procedure in $\S$~\ref{sec:simulations}.  We discuss how our model impacts rotation period distributions and the implication for gyrochronology methods in $\S$~\ref{sec:protDist} and $\S$~\ref{sec:gyro}, apply our model to the \kepler field in $\S$~\ref{sec:kepler}, and discuss our results' implications in $\S$~\ref{sec:discussion}.

@@

In Fig.~\ref{fig:stellarExample}, we plot the evolution of $R$, $r_g$, and P$_{rot}$ for 0.2 M$_{\odot}$, 0.7 M$_{\odot}$, and 1 M$_{\odot}$ mass stars, representing an M, K , and G dwarf, respectively, computed according to our stellar evolution model, \stellar, using both magnetic braking models. We assume all stars have an initial P$_{rot} = 1$ d and have an initial age of 5 Myr. All stars' radii contract along the pre-main sequence, spinning the stars up (right panel). Once the stars reach the main sequence, their structure changes slowly, allowing magnetic braking to dominate the stellar angular momentum evolution, significantly spinning-down the stars over long timescales. Both magnetic braking models produce qualitatively similar P$_{rot}$ evolution, with the \citet{Reiners2012} predicting marginally stronger torques. The $r_g$ evolution noticeably differs between the stars as the late M dwarf's (green) $r_g$ varies little as it remains full convective, while the K and G dwarf grow a radiative core while on the pre-main sequence, decreasing $r_g$ until both reach the main sequence.

@@

Here we examine the P$_{rot}$ evolution of 1 M$_{\odot} - 1$ M$_{\odot}$ binaries on initially circular orbits with a focus on systems with weak tidal torques, i.e. long P$_{orb}$ and large $Q$ or small $\tau$, to identify the boundary between evolution dominated by tides or magnetic braking.

\subsubsection{Using the \citet{Matt2015} Magnetic Braking Model}

In Fig.~\ref{fig:eqPer}, we plot the evolution of P$_{rot}$, normalized by P$_{eq}$, and its time derivative for P$_{orb} \in [5,60]$ d modeled using both the CPL (solid line, $Q=10^6$) and CTL (dashed line, $\tau = 0.1$ s) formalisms. Both models predict that binaries with P$_{orb} < 10$ d will tidally-lock within 100 Myr, in agreement with observations \citep{Meibom2005} and previous theoretical work \citep{Zahn1989}. The CPL model predicts that each binary tidally-locks, even out to P$_{orb} = 60$ d, indicating that tidal locking is not necessarily restricted to short P$_{orb}$ systems.

@@

\subsubsection{Using the \citet{Reiners2012} Magnetic Braking Model}

The mechanism for producing subsynchronous rotation is not limited to the \citet{Matt2015} magnetic braking model, but rather is a generic outcome of the coupling of magnetic braking and weak tidal torques. As demonstrated above, strong tidal torques under the CPL model tidally-lock each binary out to P$_{orb} = 60$ d. The CTL model, however, can only tidally-lock the P$_{orb} = 5$ d binary as the stronger magnetic braking under the \citet{Reiners2012} model forces the other binaries into subsynchronous rotation, similar to the \citet{Matt2015} simulations. Curiously, all CTL subsynchronous rotators seem to converge to P$_{rot}/$P$_{eq} \approx 1.1$.  In this state, tidal torques nearly balance magnetic braking, producing a long-lasting population of subsynchronous rotators, similar to what was observed above.  In the right panel of Fig.~\ref{fig:eqPerReiners}, we see that the net P$_{rot}$ derivatives for these binaries oscillate about 0, but are slightly negative due to the mechanism described above. This slight negative derivative is effectively negligible, however, as the torque balance will likely hold the binaries in subsynchronous rotation for 10s of Gyrs.  We note that in other exploratory simulations with larger $\tau$, we find that the 3 part mechanism described above holds, and tides win out over magnetic braking in the long-run, driving P$_{rot}$ back towards P$_{eq}$.

\begin{figure*}
	\includegraphics[width=\textwidth]{../Plots/eqPerTwoPanelReiners.pdf}
   \caption{Same format as Fig.~\ref{fig:eqPer}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
    \label{fig:eqPerReiners}%
\end{figure*}

\subsubsection{Subsynchronous Rotation at Short P$_{orb}$}

@@

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/protDist.pdf}
   \caption{P$_{rot}$ as a function of stellar mass and age according to our CPL (left), CTL (left center), and single star (right center) simulations integrated to system ages uniformly sampled over $1-7$ Gyr using the \citet{Matt2015} magnetic braking model. For each case, we only plot 2,500 systems for clarity but account for all systems when computing the marginalized distributions. Right: The P$_{rot}$ distribution for each case, marginalized over stellar mass.}%
    \label{fig:protDist}%
\end{figure*}

@@

Similar to the \citet{Matt2015} simulations discussed above, the single star simulations display a clear age-dependence of P$_{rot}$. The \citet{Reiners2012} model's magnetic braking transition at $M = 0.35$ M$_{\odot}$ produces a sharp cutoff in the P$_{rot} - M$ distribution between late M dwarfs and more massive stars.  More massive stars ($M \geq 0.35$ M$_{\odot}$) follow a smooth P$_{rot} -$ age sequence, while the late M dwarfs can remain rapidly rotating if they are young and in the saturated regime or extremely slowly rotating if the are older than about 4 Gyr. As in the \citet{Matt2015} model, the only rapid rotators (P$_{rot} \lsim 20$ d) are very young stars with ages $\lsim 2$ Gyr. The presence of extremely slowly rotating late M dwarfs, absent in the predictions of the \citet{Matt2015} model but seemingly present in nature \citep[e.g.][]{Newton2018}, could be used as an observational discriminant between the \citet{Matt2015} and \citet{Reiners2012} model, but a full investigation of which magnetic braking model accurately describes angular momentum loss in low mass stars is beyond the scope of this work.

\begin{figure*}[t]
	\includegraphics[width=\textwidth]{../Plots/protDistReiners.pdf}
   \caption{Same format as Fig.~\ref{fig:protDist}, but evolved using the \citet{Reiners2012} magnetic braking model.}%
    \label{fig:protDistReiners}%
\end{figure*}

The general trends found in the coupled stellar-tidal simulations using the \citet{Matt2015} model persist in the simulations evolved using the \citet{Reiners2012} model, demonstrating that these findings are not magnetic braking model-dependant, but rather a generic result of the interaction between magnetic braking and tidal torques in low-mass binaries. Both the CPL and CTL simulations display a clear single-star magnetic braking-dominated sequence, but as before, tidally-locked binaries appear across a wide range of P$_{rot}$ and primary star masses. Both the CPL and CTL model tidally-lock many short P$_{orb}$ stars into rapid rotation across all masses, with both models predicting $23\%$ of stars have P$_{rot} < 20$ d, compared to $11\%$ of single-stars, mostly for late M dwarfs in the latter case. As shown above, the CTL model predicts that binaries can strongly tidally-interact out to long P$_{orb}$, seen out to P$_{rot} = 80$ d, while the CPL model can tidally-lock stars out to P$_{orb} \approx 100$ d. These strong tidal interactions decouple P$_{rot}$ from age in binaries, mirroring the behavior discussed above, especially for rapid rotators. The median ages and $68\%$ interval for fast rotators (P$_{rot} \lsim 20$ d) are $2.6^{+2.8}_{-1.3}$ Gyr and $2.7^{+2.7}_{-1.4}$ Gyr according to the CPL and CTL models, respectfully, compared to much younger single stars with ages of $1.7^{+0.9}_{-0.5}$ Gyr, similar to the age distributions for rapid rotators evolved using the \citet{Matt2015} model.

@@

Tidal torques pose a fundamental problem for inferring ages of stars via gyrochronlogy. Regardless of the choice of equilibrium tidal model or magnetic braking model, stellar binaries readily tidally-lock, or at least strongly tidally-interact, across a wide range of P$_{orb}$ and primary star masses, decoupling P$_{rot}$ from age. For example, if one observed a rapidly rotating star with P$_{rot} \lsim 20$ d, gyrochronology models would predict ages $\lsim 2$ Gyr, e.g. the right panel of Fig.~\ref{fig:protDist} or Fig.~\ref{fig:protDistReiners}.  If that system is actually an unresolved binary, it would likely be tidally-locked, decoupling P$_{rot}$ from age, causing the predictions of gyrochronology models to fail. This effect is most likely to manifest in rapid rotators (P$_{rot} < 20$ d), but persists across all P$_{rot}$ up to 100 d, producing a significant contaminating signal.

In general, it is difficult to accurately determine if a source is single star or a stellar binary via longterm photometric monitoring as only a small fraction of stars in binaries will occult one another, exacerbating this issue. Observations of the binarity of field stars by \citet{Raghavan2010} and \citet{Duchene2013} indicate that roughly half of stars are in stellar binaries, with $10\%$ of these binaries having P$_{orb} \lsim 100$ d, suggesting that unless one accounts for binarity, stellar binaries will produce a significant contaminating signal in any study of stellar rotation periods and any ages inferred via gyrochronlogy are potentially subject to large systematic errors.  Moreover, this problem could be more significant than the aforementioned estimate implies as \citet{Simonian2018} found that most rapid rotators with P$_{rot} \leq 7.5$ d in the \kepler field are likely photometric binaries, suggesting that binary contamination in P$_{rot}$ studies could be widespread. We caution that any application of gyrochronlogy methods to predict ages for stars, especially those with P$_{rot} \lsim 20$ d, should rule out or account for stellar binarity, or otherwise risk deriving systematically incorrect ages.

@@

Both models can reproduce the major trends in the data, but the predictions of the CPL model fail to reproduce several key aspects.  First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all tidally-locked sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} = 1.5$.  This prediction is inconsistent with the data as no obvious spin-orbit commensurablity, aside from 1:1 synchronization, is present in the data, likely because stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.  The CPL model can produce the population of subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.  The CTL model, however, seems to excel where the CPL models fails by producing both populations. We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.

@@

The continuous supersynchronous rotation predicted by the CTL model seen in right panel of Fig.~\ref{fig:lurie7} seems to provide a much better match to the \citet{Lurie2017} data for both tests described above, providing evidence in favor of the CLT model. Moreover, it is unclear if low-mass binary stars can even lock into spin-orbit resonances, such as those discussed in \citet{Rodriguez2012}, as stellar convective envelopes lack a fixed shape \citep{Burkart2014,Lurie2017}.  There is no obvious clustering of stars near any spin-orbit resonance, aside from 1:1 spin-orbit synchronization in the \citet{Lurie2017} \kepler EB data. The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making any distinction between models conclusive.  More robust phase-resolved spectroscopic $e$ constrains would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a Gaussian process model as was done by \citet{Angus2018}.

These two tests can fail to discriminate between the CPL and CTL model, however, if the CPL model P$_{eq}$ is a continuous function of $e$, e.g. Eqn.~(\ref{eqn:cpl:eqPerCont}), as was argued by \citet{Goldreich1966b} and derived by \citet{Murray1999}.  In such a case, one would need a large number of accurate and precise measurements P$_{orb}$ and $e$, with robust uncertainties, for tidally-interacting binaries to discriminate between the CPL and CTL continuous P$_{eq}$, e.g. Eqn.~(\ref{eqn:cpl:eqPerCont}) versus Eqn.~(\ref{eqn:ctl:eqPer}).  In practice, this is likely too difficult as it requires extensive photometric and spectroscopic observations of many binaries. Since there is no obvious clustering of \kepler EBs near any spin-orbit resonance, the above two tests are likely insufficient. We offer a third observational test that provides the best method to discriminate between models.

As shown in $\S$~\ref{sec:protDist}, the detection of tidally-locked binaries with P$_{rot} \gsim 80$ d would provide strong evidence in favor of the CPL model, as the CTL model is unlikely to tidally-lock binaries past P$_{orb} \approx 20$ d and cannot tidally-lock stars beyond P$_{orb} \approx 80$ d, even with extreme $\tau$, e.g. Fig.~\ref{fig:lockedCTL}. The CPL model, however, readily tidally-locks binaries out to P$_{orb} \approx 100$ d.  We recommend observers try to measure P$_{rot}$ and $e$ in binaries out to P$_{orb} = 100$ d to test this hypothesis, but we note that detecting P$_{rot}$ for such slow rotators can be difficult due to small starspot modulation amplitudes \citep{Lurie2017,Reinhold2018}. Long term spectroscopic monitoring may be warranted in such cases.
